\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}

\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\title{\textbf{Gaussian temporal profile of gravitational decoherence\\[6pt]
from the Einstein-Langevin equation}}

\author{Will Regelmann\thanks{Corresponding author.} \quad Claude (Anthropic)\thanks{AI
assistant; contributed to mathematical development and analysis.}}

\date{February 2026}

\begin{document}
\maketitle

\begin{abstract}
Several experimental programs are approaching the sensitivity needed to test gravitational
decoherence at the mesoscopic scale. The primary theoretical target, the
Di\'osi--Penrose timescale $\tau_{DP} = \hbar/E_\Delta$, is derived either by postulating
white noise in the Newtonian gravitational potential (Di\'osi) or by invoking the
ill-definedness of the time-translation operator for superposed geometries (Penrose).
Neither route uses the Einstein equation. We ask what the Einstein-Langevin
equation---the established stochastic extension of semiclassical gravity---actually
predicts for the decoherence temporal profile of a massive body in spatial superposition.

We derive three results. (1)~The decoherence timescale is $\tau_{\text{coh}} =
(4\sqrt{2}/5)\,\hbar/E_\Delta \approx 1.13\,\tau_{DP}$, determined by the same
gravitational self-energy $E_\Delta$ as the Di\'osi--Penrose prediction. (2)~The temporal
profile is Gaussian, not exponential---a direct consequence of the noise kernel for
stationary matter being time-independent in the Newtonian limit. (3)~The profile is
material-dependent: rigid crystalline bodies give Gaussian decoherence, while soft or
amorphous bodies whose acoustic modes approach the gravitational coherence frequency
$\omega_g = E_\Delta/\hbar$ transition toward exponential. Comparing profiles across
materials of identical mass and geometry provides a discriminant absent from existing
models. The precise additional postulate that Di\'osi's model makes beyond the
Einstein-Langevin equation is identified: a white-noise temporal correlator for the
gravitational potential fluctuations.
\end{abstract}

%======================================================================
\section{Introduction}
%======================================================================

Experiments using levitated nanoparticles and matter-wave interferometry are approaching
the sensitivity needed to test gravitational decoherence at mesoscopic
masses~\cite{bose,maqro}. The principal theoretical benchmark is the Di\'osi--Penrose
timescale:
\begin{equation}
\tau_{DP} = \frac{\hbar}{E_\Delta}, \qquad
E_\Delta \equiv \frac{G}{2}\iint \frac{\Delta\rho(\mathbf{x})\,\Delta\rho(\mathbf{x}')}{|\mathbf{x}-\mathbf{x}'|}\,d^3x\,d^3x'
\label{eq:tau_dp}
\end{equation}
where $\Delta\rho = \rho_L - \rho_R$ is the mass-density difference between the two
superposed configurations. Di\'osi~\cite{diosi} derives this by postulating a white-noise
stochastic process in the Newtonian potential with a specific noise amplitude. Penrose
\cite{penrose} reaches the same timescale from the ill-definedness of the time-translation
Killing vector for superposed spacetime geometries. Di\'osi's model predicts exponential decoherence,
$|\rho_{LR}(t)| \propto e^{-t/\tau_{DP}}$; Penrose's argument identifies the same
timescale $\tau_{DP}$ but does not specify a temporal profile.

Neither derivation begins from the Einstein equation or its stochastic extension. The
Einstein-Langevin equation~\cite{hu_verdaguer} is the established framework for including
quantum stress-energy fluctuations in semiclassical gravity; it has been used to derive the
spectrum of cosmological perturbations without quantizing the metric~\cite{roura_verdaguer}
and to establish a rigorous theory of noise in curved spacetime~\cite{phillips_hu}. The
question we address is: what does this framework predict for the temporal structure of
gravitational decoherence of a massive body in spatial superposition?

The deterministic semiclassical Einstein equation alone cannot answer this question. For a
body in superposition $|\psi\rangle = (|L\rangle + |R\rangle)/\sqrt{2}$, the
stress-energy expectation value averages over branches,
$\langle\hat{T}_{\mu\nu}\rangle = \frac{1}{2}(T^{(L)}_{\mu\nu} + T^{(R)}_{\mu\nu})$,
carrying no information about the off-diagonal coherence $\rho_{LR}$. The resulting
geometry is blind to the superposition structure and cannot decohere it. The stochastic
extension is necessary.

We derive three results. First, the decoherence timescale from the Einstein-Langevin
equation is:
\begin{equation}
\tau_{\text{coh}} = \frac{4\sqrt{2}}{5}\,\frac{\hbar}{E_\Delta} \approx 1.13\,\tau_{DP}
\label{eq:tau_coh_preview}
\end{equation}
within 13\% of the DP prediction and determined by the same physical quantity $E_\Delta$.
Second, the temporal profile is Gaussian:
\begin{equation}
|\rho_{LR}(t)| = |\rho_{LR}(0)|\,e^{-t^2/\tau_{\text{coh}}^2}
\label{eq:gaussian_preview}
\end{equation}
This arises because the noise kernel for stationary matter is time-independent in the
Newtonian limit, producing static rather than white noise. Third, the profile is
material-dependent: for rigid crystalline bodies the noise is static and the profile is
Gaussian; for soft or amorphous bodies with acoustic modes near the gravitational coherence
frequency $\omega_g = E_\Delta/\hbar$, a finite noise correlation time develops and the
profile transitions toward exponential. An experimental comparison of decoherence profiles
across materials of identical mass and geometry can distinguish whether the noise is
fundamental (Di\'osi) or derived from quantum stress-energy (this work).

Section~\ref{sec:framework} recalls the Einstein-Langevin equation.
Section~\ref{sec:noise_kernel} evaluates the noise kernel for a rigid body in
superposition. Section~\ref{sec:gaussian} derives the Gaussian profile and timescale.
Section~\ref{sec:dp_comparison} identifies the precise additional postulate in Di\'osi's
model. Section~\ref{sec:predictions} states the experimental predictions.
Section~\ref{sec:discussion} discusses implications and limitations.

%======================================================================
\section{The Einstein-Langevin framework}
\label{sec:framework}
%======================================================================

The Einstein-Langevin equation is the leading-order stochastic correction to the
semiclassical Einstein equation, accounting for quantum fluctuations of the matter
stress-energy~\cite{hu_verdaguer}:
\begin{equation}
G_{\mu\nu}[g] = 8\pi G\!\left(\langle\hat{T}_{\mu\nu}\rangle_g + \xi_{\mu\nu}\right)
\label{eq:einstein_langevin}
\end{equation}
Here $G_{\mu\nu}[g]$ is the Einstein tensor, $\langle\hat{T}_{\mu\nu}\rangle_g$ is the
renormalized expectation value of the stress-energy tensor in the quantum state on the
background geometry $g$, and $\xi_{\mu\nu}$ is a real Gaussian stochastic tensor field
representing quantum stress-energy fluctuations.

The stochastic source $\xi_{\mu\nu}$ has vanishing mean,
$\langle\xi_{\mu\nu}(x)\rangle_s = 0$, and two-point function given by the noise kernel:
\begin{equation}
\langle\xi_{\mu\nu}(x)\,\xi_{\alpha\beta}(x')\rangle_s
= N_{\mu\nu\alpha\beta}(x,x')
\equiv \tfrac{1}{2}\langle\{\delta\hat{T}_{\mu\nu}(x),\,\delta\hat{T}_{\alpha\beta}(x')\}\rangle
\label{eq:noise_kernel_def}
\end{equation}
where $\delta\hat{T}_{\mu\nu} \equiv \hat{T}_{\mu\nu} - \langle\hat{T}_{\mu\nu}\rangle$
is the stress-energy fluctuation operator and $\langle\cdot\rangle$ denotes the quantum
expectation value in the physical state. In the Newtonian (non-relativistic, weak-field)
limit, the linearized metric perturbation satisfies the stochastic Poisson equation:
\begin{equation}
\nabla^2 \delta\Phi(\mathbf{x},t) = 4\pi G\,\xi_{00}(\mathbf{x},t)/c^2
\label{eq:poisson_stochastic}
\end{equation}
with retardation corrections of order $v^2/c^2$.

%======================================================================
\section{Noise kernel for a superposed rigid body}
\label{sec:noise_kernel}
%======================================================================

\subsection{Evaluation}

Consider a rigid body of mass $m$ placed in a center-of-mass superposition
$|\psi\rangle = (|L\rangle + |R\rangle)/\sqrt{2}$, with branch densities
$\rho_L(\mathbf{x}) = \rho(\mathbf{x}-\mathbf{x}_L)$ and
$\rho_R(\mathbf{x}) = \rho(\mathbf{x}-\mathbf{x}_R)$, separated by
$d = |\mathbf{x}_L - \mathbf{x}_R|$. For non-relativistic matter,
$\hat{T}_{00}(\mathbf{x}) = c^4\rho(\mathbf{x}-\hat{\mathbf{X}}_{\text{CM}})$, giving:
\begin{align}
\langle\hat{T}_{00}(\mathbf{x})\rangle &=
\tfrac{c^4}{2}\!\left[\rho_L(\mathbf{x}) + \rho_R(\mathbf{x})\right] \\
\langle\hat{T}_{00}(\mathbf{x})\hat{T}_{00}(\mathbf{x}')\rangle &=
\tfrac{c^8}{2}\!\left[\rho_L(\mathbf{x})\rho_L(\mathbf{x}')
+ \rho_R(\mathbf{x})\rho_R(\mathbf{x}')\right]
\end{align}
The cross terms $\rho_L(\mathbf{x})\rho_R(\mathbf{x}')$ and
$\rho_R(\mathbf{x})\rho_L(\mathbf{x}')$ vanish because the branches are orthogonal,
$\langle L|R\rangle = 0$: the operator
$\hat{T}_{00}(\mathbf{x})\hat{T}_{00}(\mathbf{x}')$ connects $|L\rangle$ to $|L\rangle$
and $|R\rangle$ to $|R\rangle$ (each branch has definite center-of-mass position), so only
the diagonal matrix elements survive.
Subtracting, the noise kernel~\eqref{eq:noise_kernel_def} evaluates to:
\begin{equation}
N_{0000}(\mathbf{x},\mathbf{x}') = \frac{c^8}{4}\,\Delta\rho(\mathbf{x})\,\Delta\rho(\mathbf{x}')
\label{eq:N0000}
\end{equation}
where $\Delta\rho \equiv \rho_L - \rho_R$. The noise kernel factorizes as the outer
product of the branch-difference density with itself.

\subsection{Physical interpretation}

The noise is not a quantum vacuum fluctuation of matter fields. It is the classical
statistical variance of the mass distribution arising from the superposition: the mass is
in two places, and this uncertainty generates metric fluctuations. Quantum vacuum fluctuations of $\hat{T}_{00}$
within a single branch are suppressed relative to eq.~\eqref{eq:N0000} by factors of
order $(\ell_P/\sigma)^2 \sim G\hbar/(c^3\sigma^2)$~\cite{phillips_hu}, where $\sigma$
is the body size. For a nanodiamond ($\sigma \sim 100\,\text{nm}$) this ratio is
$\sim 10^{-57}$: entirely negligible.

\subsection{Time-independence in the Newtonian limit}

For a stationary superposition---one in which wavepacket spreading is negligible on the
decoherence timescale---the mass distributions $\rho_L$ and $\rho_R$ are
time-independent. Consequently the noise kernel is static:
\begin{equation}
N_{0000}(\mathbf{x},t;\,\mathbf{x}',t') = \frac{c^8}{4}\,\Delta\rho(\mathbf{x})\,\Delta\rho(\mathbf{x}')
\label{eq:N0000_static}
\end{equation}
Retardation corrections are $\mathcal{O}(v^2/c^2)$ and negligible. The static character
of eq.~\eqref{eq:N0000_static} is the key structural feature that distinguishes the
Einstein-Langevin noise from the white noise postulated by Di\'osi.

The stationary approximation holds when the wavepacket spreading time $Md^2/\hbar$ exceeds
the decoherence timescale. For a microdiamond ($m \sim 10^{-14}$~kg, $d \sim 250\,\mu$m),
this ratio is $\sim 6\times10^9$~s versus $\sim 1$~ms---a margin of $10^{12}$.

%======================================================================
\section{Gaussian decoherence timescale}
\label{sec:gaussian}
%======================================================================

\subsection{Decoherence integral}

The stochastic Newtonian potential at branch location $\mathbf{x}_{L,R}$ is:
\begin{equation}
\delta\Phi(\mathbf{x}) = -\frac{G}{c^2}\int \frac{\xi_{00}(\mathbf{x}')}{|\mathbf{x}-\mathbf{x}'|}\,d^3x'
\end{equation}
The branch-differential stochastic force $\delta V_\Delta(t) \equiv m[\delta\Phi(\mathbf{x}_L,t) - \delta\Phi(\mathbf{x}_R,t)]$ dephases the off-diagonal element of the density matrix. Averaging over stochastic realizations~\cite{hu_verdaguer}:
\begin{equation}
|\rho_{LR}(t)| = |\rho_{LR}(0)|\exp\!\left(-\frac{1}{2\hbar^2}\int_0^t\!\!\int_0^t
C_V(t'-t'')\,dt'\,dt''\right)
\label{eq:decoherence_integral}
\end{equation}
where $C_V(\tau) \equiv \langle\delta V_\Delta(t)\,\delta V_\Delta(t+\tau)\rangle_s$.

\subsection{Static noise gives Gaussian suppression}

From eqs.~\eqref{eq:N0000_static} and~\eqref{eq:poisson_stochastic}, the correlator
inherits time-independence: $C_V(\tau) = C_V(0)$ for all $\tau$. The double integral in
eq.~\eqref{eq:decoherence_integral} then evaluates to $C_V(0)\,t^2$, giving:
\begin{equation}
|\rho_{LR}(t)| = |\rho_{LR}(0)|\,\exp\!\left(-\frac{C_V(0)}{2\hbar^2}\,t^2\right)
\label{eq:gaussian_suppression}
\end{equation}
This is the direct mathematical consequence of static noise. In the Markovian limit, a
finite noise correlation time $\tau_c$ gives $\int\int C_V dt'dt'' \approx 2C_V(0)\tau_c
t$, yielding exponential decoherence at rate $\Gamma = 2C_V(0)\tau_c/\hbar^2$. In the
static limit ($\tau_c\to\infty$ for stationary matter), the exponential rate vanishes and
the profile is Gaussian with timescale $\tau_{\text{coh}} = \hbar\sqrt{2}/\sqrt{C_V(0)}$.

\subsection{Evaluation of $C_V(0)$ for a uniform sphere}

Propagating the factored noise kernel $N_{0000}\propto\Delta\rho\,\Delta\rho'$
through the Poisson Green's function, the stochastic-potential two-point function
factorizes as $\langle\delta\Phi(\mathbf{x})\,\delta\Phi(\mathbf{y})\rangle_s \propto U(\mathbf{x})\,U(\mathbf{y})$
where $U(\mathbf{x})\equiv \int\Delta\rho(\mathbf{z})/|\mathbf{x}-\mathbf{z}|\,d^3z$.
Consequently $\langle[\delta\Phi(\mathbf{x}_L)-\delta\Phi(\mathbf{x}_R)]^2\rangle_s
\propto [U(\mathbf{x}_L)-U(\mathbf{x}_R)]^2$, and assembling the overall prefactor:
\begin{equation}
C_V(0) = \frac{G^2 m^2}{4}\!\left[U(\mathbf{x}_L) - U(\mathbf{x}_R)\right]^2, \qquad
U(\mathbf{x}) \equiv \int\frac{\Delta\rho(\mathbf{z})}{|\mathbf{x}-\mathbf{z}|}\,d^3z
\label{eq:CV0}
\end{equation}
For a uniform sphere of radius $R$ with $d > 2R$ (non-overlapping branches), Newton's
shell theorem gives\footnote{The potential
$\int\rho(\mathbf{z})/|\mathbf{x}_L - \mathbf{z}|\,d^3z$ of a uniform sphere of mass $m$
and radius $R$, evaluated at its own center, equals $3m/(2R)$: in spherical coordinates,
$\int_0^R \rho_0 \cdot 4\pi r^2/r\,dr = 2\pi\rho_0 R^2 = 3m/(2R)$.  Since
$\Delta\rho = \rho_L - \rho_R$ and the branches do not overlap,
$U(\mathbf{x}_L) = 3m/(2R) - m/d$ (self-potential of the $L$-sphere at its center, minus
the external potential of the $R$-sphere at distance $d$). By symmetry
$U(\mathbf{x}_R) = -3m/(2R) + m/d$, giving the stated result.}:
\begin{equation}
U(\mathbf{x}_L) - U(\mathbf{x}_R) = m\!\left(\frac{3}{R} - \frac{2}{d}\right)
\end{equation}
Substituting into eq.~\eqref{eq:CV0} and recognizing
$E_\Delta = G(6m^2/5R - m^2/d)$~(eq.~\ref{eq:tau_dp}):
\begin{equation}
C_V(0) = \frac{25}{16}\,E_\Delta^2\left(1 + \mathcal{O}(R/d)\right)
\label{eq:CV0_sphere}
\end{equation}
\noindent
The leading correction\footnote{Writing $\varepsilon = R/d$ and
$f(\varepsilon) \equiv C_V(0)/E_\Delta^2 = (3-2\varepsilon)^2/[4(6/5-\varepsilon)^2]$,
one finds $f(0) = 25/16$ and $(\mathrm{d}\ln f/\mathrm{d}\varepsilon)|_{\varepsilon=0} = 1/3$, giving
$C_V(0) = \tfrac{25}{16}E_\Delta^2(1 + R/(3d) + \mathcal{O}(R^2/d^2))$ and
$\tau_{\text{coh}} = (4\sqrt{2}/5)(\hbar/E_\Delta)(1 - R/(6d) + \mathcal{O}(R^2/d^2))$.}
is positive in $C_V(0)$ and negative in $\tau_{\text{coh}}$.
The $G^2$ factor in $C_V(0)$ arises from the two-step chain: stress-energy noise converts
to metric perturbation (one factor of $G$), which converts to potential (a second). For
the overlapping regime ($d < 2R$), $E_\Delta$ must be computed via the Fourier-space
integral:
\begin{equation}
E_\Delta = \frac{2Gm^2}{\pi}\int_0^\infty|\tilde{F}(kR)|^2\!\left(1-\frac{\sin kd}{kd}\right)dk, \quad
\tilde{F}(kR) = \frac{3[\sin kR - kR\cos kR]}{(kR)^3}
\label{eq:E_delta_fourier}
\end{equation}

\subsection{Coherence timescale}

\begin{equation}
\tau_{\text{coh}} = \frac{\hbar\sqrt{2}}{\sqrt{C_V(0)}}
= \frac{4\sqrt{2}}{5}\,\frac{\hbar}{E_\Delta}\left(1 + \mathcal{O}(R/d)\right)
\approx 1.13\,\tau_{DP} \quad (d \gg R)
\label{eq:tau_coh}
\end{equation}
The $G^2$ in $C_V(0)$ yields $G^{-1}$ scaling in $\tau_{\text{coh}}$ via the square root,
matching the $G$-scaling of $\tau_{DP}$. Numerical values for three experimental platforms
are given in Table~\ref{tab:platforms}.

\begin{table}[h]
\centering
\begin{tabular}{lcccccc}
\toprule
Platform & $m$ & $R$ & $d$ & $\tau_{DP}$ & $\tau_{\text{coh}}$ & Profile \\
\midrule
Microdiamond (BMV)~\cite{bose} & $10^{-14}$~kg & 90~nm & 250~$\mu$m & 1.2~ms & 1.3~ms & Gaussian \\
Levitated nanosphere & $5\times10^{-18}$~kg & 100~nm & 500~nm & 6300~s & 7100~s & Gaussian \\
MAQRO test mass~\cite{maqro} & $1.7\times10^{-17}$~kg & 200~nm & 100~nm & 10700~s & 12100~s & Gaussian \\
\bottomrule
\end{tabular}
\caption{Coherence timescales for three experimental platforms. The MAQRO entry uses
eq.~\eqref{eq:E_delta_fourier} (overlapping-branch regime, $d/R = 0.5$); in this
regime the prefactor $4\sqrt{2}/5$ in eq.~\eqref{eq:tau_coh} does not apply and
$\tau_{\text{coh}}$ is obtained numerically from $\tau_{\text{coh}} = \hbar\sqrt{2}/\sqrt{C_V(0)}$
with $C_V(0)$ evaluated via eq.~\eqref{eq:E_delta_fourier}. All three platforms
satisfy $\omega_{\text{phonon}}/\omega_g \gg 1$ and are in the rigid-body Gaussian regime.}
\label{tab:platforms}
\end{table}

%======================================================================
\section{Relation to the Di\'osi--Penrose prediction}
\label{sec:dp_comparison}
%======================================================================

\subsection{Same timescale}

Both the Einstein-Langevin derivation and the Di\'osi--Penrose prediction give a timescale
set by $E_\Delta$. This is not coincidental: the bilinear structure $\langle\Delta\rho,\Delta\rho'\rangle_{1/r}$
of the noise kernel selects $E_\Delta$ as the natural energy scale in the Newtonian limit. The specific prefactor
$4\sqrt{2}/5 \approx 1.13$ in eq.~\eqref{eq:tau_coh} comes from the uniform-sphere
geometry and could differ for other mass distributions.

\subsection{Different temporal correlator: the extra postulate}

The profiles differ because the noise correlators differ. The Einstein-Langevin equation
gives a static correlator (eq.~\ref{eq:N0000_static}), which propagates to:
\begin{equation}
C_V^{\text{EL}}(\tau) = C_V(0) \quad \forall\,\tau
\label{eq:corr_el}
\end{equation}
Di\'osi's model postulates a white-noise correlator:
\begin{equation}
C_V^{\text{D}}(\tau) = 2C_V(0)\,\tau_c\,\delta(\tau)
\label{eq:corr_diosi}
\end{equation}
with $\tau_c$ chosen so that $\Gamma = C_V(0)\tau_c/\hbar^2 = \hbar/\tau_{DP}^2$ reproduces
the DP rate. The spatial structure and equal-time amplitude $C_V(0)$ are identical in both
cases; the difference is entirely in the temporal correlator. Di\'osi's extra postulate,
beyond what the Einstein-Langevin equation specifies, is the Markovian (white-noise)
character of the temporal correlations.

A summary comparison with the Kafri--Taylor--Milburn model~\cite{kafri} is given in
Table~\ref{tab:models}.

\begin{table}[h]
\centering
\begin{tabular}{lllll}
\toprule
Model & Timescale & $G$-scaling & Profile & Noise temporal correlator \\
\midrule
This work & $\hbar/E_\Delta$ & $1/G$ & Gaussian & Static (derived) \\
Di\'osi--Penrose & $\hbar/E_\Delta$ & $1/G$ & Exponential & White (postulated) \\
Kafri--Taylor--Milburn & $\hbar D^5/(G^2 m^3)$ & $1/G^2$ & Exponential & White (postulated) \\
\bottomrule
\end{tabular}
\caption{Comparison of gravitational decoherence models. This work and Di\'osi--Penrose
share the same timescale scaling; the discriminating prediction is the temporal profile.}
\label{tab:models}
\end{table}

%======================================================================
\section{Experimental predictions}
\label{sec:predictions}
%======================================================================

\subsection{Profile discrimination}

The maximum fractional difference between the predicted profiles occurs near
$t \approx \tau_{DP}/2$. Using $|\rho_{LR}^{\text{Gauss}}| = e^{-t^2/\tau_{\text{coh}}^2}$
with $\tau_{\text{coh}} = 1.13\,\tau_{DP}$, and
$|\rho_{LR}^{\text{exp}}| = e^{-t/\tau_{DP}}$:
\begin{equation}
\max_t\frac{|\rho_{LR}^{\text{Gauss}}| - |\rho_{LR}^{\text{exp}}|}{|\rho_{LR}(0)|}
\approx 0.22
\label{eq:delta_V}
\end{equation}
This maximum occurs at $t \approx 0.48\,\tau_{DP} \approx 0.43\,\tau_{\text{coh}}$,
where the Gaussian has decayed by $\sim18\%$ and the exponential by $\sim40\%$.
In the limiting approximation $\tau_{\text{coh}} \approx \tau_{DP} \equiv \tau$,
the difference at $t = \tau/2$ simplifies to $e^{-1/4} - e^{-1/2} \approx 0.17$.
To resolve the actual 0.22 difference at $3\sigma$ requires per-time-point visibility
uncertainty $\sigma_V < 0.07$. For shot-noise-limited measurements with fringe contrast
$C \approx 0.5$ this demands $N_{\text{run}} \gtrsim 1000$ shots per time point. Across
eight time points spanning $t \in [0.1\tau_{DP}, 3\tau_{DP}]$, the total is
$\sim 8{,}000$ experimental shots.

The profile discrimination requires that environmental decoherence $\Gamma_{\text{env}}$
be characterized to within a factor of $\sim 2$ (so that $\delta\Gamma_{\text{env}}\cdot
\tau_{\text{coh}} \lesssim 0.1$). This is achievable by comparing data at different masses
or between materials with the same environmental coupling but different $E_\Delta$.

\subsection{Material-dependent profile transition}
\label{sec:material}

The static-noise result~\eqref{eq:N0000_static} holds when the body's lowest acoustic
phonon frequency $\omega_{\text{phonon}}$ far exceeds the gravitational coherence frequency
$\omega_g = E_\Delta/\hbar$. For nanoscale diamond, $\omega_{\text{phonon}} \sim c_s/L
\sim 10^{10}$~Hz while $\omega_g \sim 10^3$~rad/s for $m \sim 10^{-14}$~kg: the ratio
$\omega_{\text{phonon}}/\omega_g \sim 10^7$ confirms that the rigid-body Gaussian result
applies to all current experimental targets.

When $\omega_{\text{phonon}} \sim \omega_g$, internal dynamics introduce a finite noise
correlation time $\tau_c \sim \omega_{\text{phonon}}^{-1}$, and the decoherence exponent
crosses over:
\begin{equation}
\int_0^t\!\!\int_0^t C_V(t'-t'')\,dt'dt'' \approx
\begin{cases}
C_V(0)\,t^2 & t \ll \tau_c \\
2\,C_V(0)\,\tau_c\,t & t \gg \tau_c
\end{cases}
\label{eq:crossover}
\end{equation}
This predicts qualitatively different profiles for bodies of identical mass and geometry
but different material rigidity:
\begin{itemize}
\item \textbf{Rigid crystalline solids} (diamond, silicon, sapphire;
$\omega_{\text{phonon}} \sim 10^{10}$--$10^{12}$~Hz): Gaussian profile throughout the
accessible coherence-frequency range.
\item \textbf{Soft or amorphous materials} (polymers, aerogels, amorphous metals;
$\omega_{\text{phonon}} \sim 10^4$--$10^6$~Hz): Intermediate or exponential profile when
$\omega_g \sim \omega_{\text{phonon}}$.
\end{itemize}
The Di\'osi model and Penrose collapse model predict exponential decoherence regardless of
material. Observing profile differences across materials at fixed mass and geometry would
directly probe whether the noise originates in quantum stress-energy (this work) or is a
fundamental background process.

\subsection{BMV entanglement}

In the Bose--Marletto--Vedral experiment~\cite{bose,marletto_vedral}, gravitational
entanglement generation between two superposed masses would signal quantum gravity. In the
Einstein-Langevin framework, gravity mediates via a classical stochastic channel. For a
bipartite product state $\rho_1\otimes\rho_2$, the metric perturbation factorizes as
$h_{\mu\nu} = h^{(1)}_{\mu\nu} + h^{(2)}_{\mu\nu}$, yielding a c-number interaction
potential that acts as a classical common-cause channel. Because entanglement cannot
be generated by local operations on a separable state coupled only through a classical
stochastic field~\cite{oppenheim,marletto_classical}, the prediction is no entanglement at the quantum
gravity rate; observation of such entanglement would falsify the framework.

%======================================================================
\section{Discussion}
\label{sec:discussion}
%======================================================================

\textbf{What the derivation assumes.} The results rest on four standard assumptions within
the Hu-Verdaguer stochastic gravity program: (i)~the Einstein-Langevin equation describes
metric fluctuations sourced by quantum stress-energy; (ii)~the Newtonian limit applies;
(iii)~the superposition is stationary on the decoherence timescale; (iv)~leading-order
perturbation theory (backreaction of decoherence on the noise kernel is neglected).
Backreaction becomes significant when the compactness parameter $Gm/(c^2 R)$ approaches
unity, i.e., when the body's gravitational radius is comparable to its size. For all
laboratory masses this ratio is negligible: $Gm/(c^2 R) \sim 10^{-35}$ for a
microdiamond. Perturbation theory breaks down only for compact objects with
$R \sim Gm/c^2$, far outside the experimental regime considered here.
No physics beyond the semiclassical Einstein equation is introduced.

\textbf{State-dependent noise.} The noise kernel~\eqref{eq:N0000} vanishes when
$\Delta\rho = 0$ (no superposition). The metric fluctuations that decohere the
superposition are generated by the superposition itself, not by a background stochastic
process. This is physically consistent and distinguishes the present framework
from Di\'osi's model, where the noise is a fundamental property of spacetime.

\textbf{Limitations.} The Gaussian profile is a leading-order prediction; retardation
corrections are $\mathcal{O}(v^2/c^2)$. The material-dependent crossover
(eq.~\ref{eq:crossover}) requires a more detailed model of the phonon-spectrum contribution
to the noise kernel, which we identify as an open problem. The BMV null prediction is
necessary but not sufficient; the discriminating positive prediction is the temporal
profile and its material dependence.

\textbf{Companion paper.} The semiclassical Einstein equation underlying the
Einstein-Langevin formalism is itself a fixed-point condition: geometry and quantum state
must be mutually self-consistent. The existence of self-consistent solutions---established
exactly for the Starobinsky inflation model and conditionally in general via the Schauder
fixed-point theorem---is treated in a companion paper~\cite{fixed_point}.

%======================================================================
\begin{thebibliography}{99}

\bibitem{bose} S.~Bose \textit{et al.}, Spin entanglement witness for quantum gravity,
\textit{Phys.\ Rev.\ Lett.}\ \textbf{119}, 240401 (2017).

\bibitem{maqro} R.~Kaltenbaek \textit{et al.}, Macroscopic quantum resonators (MAQRO),
\textit{Exp.\ Astron.}\ \textbf{34}, 123 (2012).

\bibitem{diosi} L.~Di\'osi, A universal master equation for the gravitational violation
of quantum mechanics, \textit{Phys.\ Lett.\ A}\ \textbf{120}, 377 (1987).

\bibitem{penrose} R.~Penrose, On gravity's role in quantum state reduction,
\textit{Gen.\ Rel.\ Grav.}\ \textbf{28}, 581 (1996).

\bibitem{hu_verdaguer} B.~L.~Hu and E.~Verdaguer, Stochastic gravity: Theory and
applications, \textit{Living Rev.\ Relativ.}\ \textbf{11}, 3 (2008).

\bibitem{roura_verdaguer} A.~Roura and E.~Verdaguer, Cosmological perturbations from
stochastic gravity, \textit{Phys.\ Rev.\ D}\ \textbf{78}, 064010 (2008).

\bibitem{phillips_hu} N.~G.~Phillips and B.~L.~Hu, Noise kernel in stochastic gravity and
stress energy bitensor, \textit{Phys.\ Rev.\ D}\ \textbf{63}, 104001 (2001).

\bibitem{marletto_vedral} C.~Marletto and V.~Vedral, Gravitationally induced entanglement
between two massive particles is sufficient evidence of quantum effects in gravity,
\textit{Phys.\ Rev.\ Lett.}\ \textbf{119}, 240402 (2017).

\bibitem{kafri} D.~Kafri, J.~M.~Taylor, and G.~J.~Milburn, A classical channel model for
gravitational decoherence, \textit{New J.\ Phys.}\ \textbf{16}, 065020 (2014).

\bibitem{oppenheim} J.~Oppenheim, A postquantum theory of classical gravity?,
\textit{Phys.\ Rev.\ X}\ \textbf{13}, 041040 (2023).

\bibitem{marletto_classical} C.~Marletto, J.~Oppenheim, V.~Vedral, and E.~Wilson,
Classical gravity cannot mediate entanglement,
preprint arXiv:2511.07348 (2025).

\bibitem{fixed_point} W.~Regelmann and Claude (Anthropic), Fixed-point existence of
self-consistent semiclassical gravity (2026). Companion paper.

\end{thebibliography}

\end{document}
