\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}

\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{axiom}{Axiom}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{conjecture}{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{remark}
\newtheorem{remark}{Remark}

\title{\textbf{Co-emergence of mass, signature, time, and Hilbert space}\\[6pt]
{\large DRAFT}}

\author{Will Regelmann\thanks{Corresponding author.} \quad Claude (Anthropic)\thanks{AI
assistant; contributed to mathematical development and analysis.}}

\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
We argue that mass, Lorentzian signature, local time, and local Hilbert
space are four aspects of a single self-consistent structure: none exists
without the others. Physical mass requires hyperbolicity (Wigner 1939);
hyperbolicity provides local proper time along timelike geodesics; local
proper time enables a local Hilbert space via the Page--Wootters mechanism;
and the resulting quantum stress-energy sources gravity, closing the
self-consistency loop. We formalize this co-emergence within a four-level
hierarchy of geometric constraints on a block 4-manifold with no evolution
parameter, no background structure, and no fundamental Hilbert space.
The semiclassical Einstein equation --- the hierarchy's metric level ---
is signature-blind: self-consistent solutions exist for both Lorentzian
and Riemannian metrics, for both massive and massless field content.
Signature selection operates not at any single level but at the
cross-level constraint, where only Lorentzian metrics with massive field
content support the full hierarchy.

\bigskip\noindent\textit{Note:} Companion to ``Fixed-point existence of
self-consistent semiclassical gravity''~\cite{fixed_point}, whose results
are cited but not re-derived.
\end{abstract}

%======================================================================
\section{Introduction}
\label{sec:intro}
%======================================================================

Every formulation of fundamental physics assumes Lorentzian signature.
The distinction between one timelike and three spacelike directions---the
$(-,+,+,+)$ sign pattern in the metric---is an input, not a derivation.
If the universe is a four-dimensional geometric block satisfying
constraints rather than a state evolving in time, this distinction should
be a consequence of the constraints, not a separate postulate.

The most natural route to deriving signature would be topological: the
intersection form $Q_\mathcal{M}$ of a 4-manifold has a type $(b^+, b^-)$
that mirrors the $(1,3)$ split of Lorentzian signature. But this route is
obstructed. The intersection form is a global invariant on
$H_2(\mathcal{M};\mathbb{Z})$, while metric signature is a pointwise
property of the tangent bundle; no theorem connects them. Worse, closed
simply-connected 4-manifolds with nontrivial intersection forms have
Euler characteristic $\chi \geq 3$, while a closed manifold admits a
Lorentzian metric only if $\chi = 0$~\cite{geroch}. The manifold class
where the Freedman--Donaldson machinery~\cite{freedman,donaldson} is
strongest cannot carry a Lorentzian metric at all.

The semiclassical Einstein equation (SCE) provides no help either. The
companion paper~\cite{fixed_point} establishes that the perturbative
self-consistency map has contraction constant $\kappa \sim (m/M_P)^2$
regardless of metric signature. The Banach contraction is
signature-blind: self-consistent solutions exist for both Lorentzian and
Riemannian metrics. The Starobinsky de Sitter solution~\cite{starobinsky}
extends this blindness to massless conformal matter. Level by level,
signature is not selected.

The foundational tension in quantum gravity is well known: general
relativity requires no preferred time parameter, while standard quantum
mechanics requires one. The Wheeler--DeWitt equation
$\hat{H}\Psi = 0$~\cite{dewitt,wheeler} is the canonical attempt at
resolution---the Hamiltonian constraint makes the wave functional
timeless. But every known formulation of $\hat{H}\Psi = 0$ introduces
hidden temporal or foliation structure at some stage of its
derivation~\cite{kuchar}. The canonical ADM approach requires a $3+1$
decomposition; the covariant path integral hides foliation dependence in
the Wick rotation, boundary conditions, and measure ambiguity; the WKB
semiclassical limit inserts time via Born--Oppenheimer
adiabaticity~\cite{halliwell,banks}.

This paper proposes that signature and time are two faces of the same
coin, and that neither can be derived in isolation. Mass, Lorentzian
signature, local time, and local Hilbert space are four aspects of a
single self-consistent structure: each requires the others, and all
emerge together from the requirement that the geometric block satisfy
constraints at every level simultaneously. We call this the
\emph{co-emergence thesis}. It is formalized as a four-level hierarchy
of geometric constraints (Section~\ref{sec:hierarchy}) whose joint
solution space is restricted to Lorentzian signature with massive field
content (Conjecture~\ref{conj:coemergence}).

The paper is organized as follows. Section~\ref{sec:axioms} states the
three axioms. Section~\ref{sec:coemergence} presents the co-emergence
thesis: four links forming a closed chain, each with its rigor level.
Section~\ref{sec:hierarchy} develops the four-level hierarchy.
Section~\ref{sec:dimensions} argues that four-dimensionality is derived,
not assumed. Section~\ref{sec:connections} connects the hierarchy to
Wheeler--DeWitt, the SCE, Page--Wootters, and Osterwalder--Schrader
reconstruction. Section~\ref{sec:open} lists priority open problems.

%======================================================================
\section{Axioms}
\label{sec:axioms}
%======================================================================

\begin{axiom}[No evolution parameter]
\label{ax:timeless}
The fundamental description contains no evolution parameter. All equations
are constraints on the structure of the four-dimensional block, not rules for
generating it. A formulation satisfies this axiom if and only if it can be
stated entirely as a constraint on a four-dimensional geometric structure with
no preferred slicing.
\end{axiom}

\begin{axiom}[No background structure]
\label{ax:nobackground}
The fundamental description contains no fixed metric, no preferred
foliation, no assumed symmetry group, and no presupposed dimensionality
or signature. The consistency constraints are purely geometric: at the
deepest level, metric structure and matter both emerge from
topology~(Conjecture).
\end{axiom}

\begin{axiom}[No fundamental Hilbert space]
\label{ax:nohilbert}
The fundamental description does not take a Hilbert space as input.
Hilbert space, if it appears, is derived---an effective description for
subsystems with limited access to the full block.
\end{axiom}

Axiom~\ref{ax:nohilbert} requires justification, since every existing
formulation of quantum gravity starts from a Hilbert space.

A Hilbert space $\mathcal{H}$ with inner product
$\langle\cdot|\cdot\rangle$ carries temporal structure in three ways,
any one of which conflicts with Axiom~\ref{ax:timeless}:

\begin{enumerate}

\item \textbf{The inner product requires a time-slice.} In quantum field
theory, $\langle\Psi_1|\Psi_2\rangle = \int \Psi_1^*[\phi]\Psi_2[\phi]
\,\mathcal{D}\phi$ integrates over field configurations $\phi$ on a spatial
hypersurface at a fixed time. Different choices of hypersurface give
inequivalent inner products and inequivalent quantum theories.

\item \textbf{Unitary evolution requires a time parameter.} The statement
$|\Psi(t_2)\rangle = U(t_2, t_1)|\Psi(t_1)\rangle$ requires a preferred
parameter $t$. Even in the ``frozen'' Wheeler--DeWitt formulation, the
Hilbert space inner product is defined on a spatial slice, smuggling in
the foliation through the back door.

\item \textbf{The Born rule requires a preferred decomposition.}
The probability $|\langle\phi|\Psi\rangle|^2$ is defined relative to a
measurement basis $\{|\phi\rangle\}$. Selecting this basis is an
observer-dependent choice that the fundamental description should not
require.

\end{enumerate}

\begin{proposition}[Hilbert space violates Axiom~\ref{ax:timeless}]
\emph{(Rigorous for (1) and (2); Sketch for (3).)}
\label{prop:nohilbert}
Any formulation of quantum gravity that takes a Hilbert space with inner
product as a fundamental input violates Axiom~\ref{ax:timeless}, because the
inner product requires a preferred spatial hypersurface (point~(1)) or a
preferred time parameter (point~(2)).
\end{proposition}

\noindent\emph{Argument.} Point~(1) follows from the standard QFT
definition of the Fock space inner product: the functional integral
$\int\mathcal{D}\phi$ is over field configurations on a Cauchy surface,
and different Cauchy surfaces give different inner products. Point~(2) is
immediate from the definition of unitary evolution. \hfill$\square$

\medskip

The conclusion is that a formulation satisfying Axiom~\ref{ax:timeless}
must not take Hilbert space as input. If Hilbert space appears at all,
it must emerge as a structure for subsystems---the content of
Section~\ref{sec:coemergence}.

\medskip

\noindent\textbf{What this paper is and is not.} This is a framework
proposal, not a completed theory. The co-emergence thesis
(Section~\ref{sec:coemergence}) rests on specific mathematical claims at
stated rigor levels. The metric level has the most rigorous support---it
is the setting of the companion paper~\cite{fixed_point}. The
topological, smooth-structure, and effective-QM levels are research
programs with clear mathematical agendas. We do not claim to have derived
the Standard Model. We claim to have identified the structure that
connects mass, signature, time, and Hilbert space, and to have made
explicit what is required to prove the connection rigorously.

%======================================================================
\section{The co-emergence thesis}
\label{sec:coemergence}
%======================================================================

Four concepts that are usually treated as independent---mass, Lorentzian
signature, local time, and local Hilbert space---form a closed logical
chain. Each link has a stated rigor level and citation. The chain reads
as a causal sequence; it is not. The four concepts are simultaneous
aspects of the self-consistent block (see Remark~\ref{rem:temporal}
below).

%----------------------------------------------------------------------
\subsection{Mass requires Lorentzian signature}
\label{sec:mass_signature}
%----------------------------------------------------------------------

\emph{(Rigorous.)}

\medskip

\noindent The distinction between physical mass and the parameter $m^2$
is categorical. On a Lorentzian manifold, the dispersion relation
$E^2 = \mathbf{p}^2 + m^2$ defines a mass shell $p_\mu p^\mu = -m^2$---a
hyperboloid in momentum space. Massive particles are irreducible
representations of the Poincar\'e group $\mathrm{ISO}(3,1)$, classified
by mass and spin (Wigner~\cite{wigner}). The mass shell, the dispersion
relation, and the particle classification all require the $(-,+,+,+)$
split.

On a Riemannian manifold, the parameter $m^2$ appears in the field
equation $(\Delta + m^2)\phi = 0$ but does not define a mass shell, a
dispersion relation, or propagating degrees of freedom. The wave operator
$\Delta + m^2$ is elliptic: its solutions decay exponentially rather than
propagate. The Euclidean group $\mathrm{ISO}(4)$ has no analogous
classification of particles by mass and spin. The parameter $m^2$ exists;
the physics of mass does not.

%----------------------------------------------------------------------
\subsection{Lorentzian signature provides local time}
\label{sec:signature_time}
%----------------------------------------------------------------------

\emph{(Rigorous; Geroch~\cite{geroch}, Hawking--Ellis~\cite{hawking_ellis}.)}

\medskip

\noindent Lorentzian signature distinguishes timelike, spacelike, and null
directions. This trichotomy is the physical content of the $(-,+,+,+)$
sign pattern: it is what makes the metric a causal structure and not merely
a distance function. Massive particles follow timelike geodesics, along
which proper time
\begin{equation}
d\tau^2 = -g_{\mu\nu}\,dx^\mu dx^\nu > 0
\label{eq:propertime}
\end{equation}
is well-defined and positive. Massless particles follow null geodesics,
along which $d\tau = 0$.

General relativity prohibits universal time: clocks at different locations
in a gravitational field tick at different rates. But mass provides
\emph{local} time: worldline-dependent clocks carried by massive objects.
Without hyperbolicity, there is no causal structure, no
timelike/spacelike distinction, and no proper time for any object.

%----------------------------------------------------------------------
\subsection{Local time enables local Hilbert space}
\label{sec:time_hilbert}
%----------------------------------------------------------------------

\emph{(Sketch.)}

\medskip

\noindent The Schr\"odinger equation $i\hbar\,\partial|\psi\rangle/\partial t
= H|\psi\rangle$ requires a time parameter $t$. The inner product
$\langle\psi|\phi\rangle$ is defined on a spatial slice at fixed $t$.
Unitarity requires that the inner product be preserved under evolution in
$t$. If there is no universal time (Section~\ref{sec:signature_time}),
there is no universal Hilbert space---the content of
Axiom~\ref{ax:nohilbert}.

But if mass provides local time, a local Hilbert space exists for massive
subsystems that carry their own clocks. The Page--Wootters
mechanism~\cite{page_wootters} makes this precise: apparent time evolution
is recovered from a globally static state by conditioning on a clock
subsystem. The clock must be massive---it must experience proper time along
a timelike worldline. Without mass, the conditioning has no temporal
reference and the mechanism fails.

This is how Axiom~\ref{ax:nohilbert} is realized: Hilbert space is not
fundamental but is derived via the Page--Wootters mechanism from mass and
Lorentzian signature.

\begin{remark}[Acknowledged gap]
\label{rem:pw_gap}
The Page--Wootters mechanism as standardly formulated presupposes a total
Hilbert space $\mathcal{H} = \mathcal{H}_{\mathrm{clock}} \otimes
\mathcal{H}_{\mathrm{system}}$ and a global Hamiltonian constraint
$\hat{H}|\Psi\rangle = 0$. This conflicts with Axiom~\ref{ax:nohilbert}:
the total Hilbert space is assumed, not derived. A fully consistent
treatment requires reformulating the Page--Wootters mechanism without a
fundamental Hilbert space---using instead the marginal of the
smooth-structure measure $\mu^*$ (Section~\ref{sec:level3}). This
reformulation is an open problem (Section~\ref{sec:open}).
\end{remark}

%----------------------------------------------------------------------
\subsection{Local Hilbert space closes the loop}
\label{sec:hilbert_mass}
%----------------------------------------------------------------------

\emph{(Sketch.)}

\medskip

\noindent The Hilbert space for massive quantum fields produces the
quantum stress-energy $\langle\hat{T}_{\mu\nu}\rangle$ that sources
gravity via the semiclassical Einstein equation
\begin{equation}
G_{\mu\nu}[g] = 8\pi G\,\langle\hat{T}_{\mu\nu}\rangle_{g,\mathrm{ren}}.
\label{eq:sce}
\end{equation}
The companion paper~\cite{fixed_point} establishes that eq.~\eqref{eq:sce}
has a self-consistent solution (Rigorous for massive fields with
$m \ll M_P$). Without mass, the only stress-energy source is the conformal
trace anomaly, which drives the Starobinsky de Sitter
solution~\cite{starobinsky}. This solution is self-consistent at the
metric level (Section~\ref{sec:level2}), but it supports no massive
particles, no local clocks, no local Hilbert space, and no effective
quantum mechanics for subsystems. It is a geometry in equilibrium with
vacuum fluctuations---consistent but physically uninhabited.

Mass makes the self-consistent geometry physically inhabited: it provides
the matter content that sources gravity, the timelike worldlines that
define local time, and the clock subsystems that enable local quantum
mechanics. The loop closes:
\begin{equation}
\text{mass} \;\longleftrightarrow\; \text{Lorentzian signature}
\;\longleftrightarrow\; \text{local time}
\;\longleftrightarrow\; \text{local Hilbert space}
\;\longleftrightarrow\; \langle\hat{T}_{\mu\nu}\rangle
\;\longleftrightarrow\; \text{gravity}
\;\longleftrightarrow\; \text{mass}
\label{eq:co_emergence}
\end{equation}
In the framework of Axiom~\ref{ax:timeless}, this is not a circularity
but a self-consistency constraint: the block must satisfy all these
conditions simultaneously.

%----------------------------------------------------------------------
\subsection{The formal statement}
\label{sec:formal_coemergence}
%----------------------------------------------------------------------

\begin{conjecture}[Co-emergence]
\label{conj:coemergence}
Cross-level self-consistency of the four-level hierarchy
(Definition~\ref{def:hierarchy}) has nontrivial solutions---solutions
supporting effective quantum mechanics for subsystems---only for
Lorentzian metric signature with massive field content. Riemannian
and/or massless solutions exist at the metric level but are physically
degenerate: they do not support the Level~3 structure.
\end{conjecture}

The evidence for Conjecture~\ref{conj:coemergence} is summarized in
Table~\ref{tab:signature}. The semiclassical Einstein equation (Level~2)
is completely signature-blind: self-consistent fixed points exist for
all four combinations of signature and field content. The selection
operates at Level~3, where only the Lorentzian-massive combination
supports local time, local Hilbert space, and effective quantum
mechanics.

\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Massive fields} & \multicolumn{2}{c}{Massless conformal} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
& Level~2 & Level~3 & Level~2 & Level~3 \\
\midrule
Lorentzian & $\checkmark$ (Rigorous) & $\checkmark$ (Sketch) & $\checkmark$ (Rigorous) & $\times$ \\
Riemannian & $\checkmark$ (Sketch) & $\times$ & $\checkmark$ (Rigorous) & $\times$ \\
\bottomrule
\end{tabular}
\caption{Signature selection operates at the cross-level constraint. Level~2
(the SCE) is signature-blind for both massive and massless fields. Only
Lorentzian signature with massive field content supports Level~3.
The Lorentzian massive Level~2 entry is Rigorous (Banach contraction,
$\kappa \sim (m/M_P)^2$, companion paper~\cite{fixed_point}); the
Riemannian massive entry is Sketch (same contraction argument applies to
elliptic operators); the massless entries use the Starobinsky de Sitter /
round $S^4$ solutions~\cite{starobinsky,bunch_davies}.}
\label{tab:signature}
\end{table}

\begin{remark}[The chain is not causal]
\label{rem:temporal}
The co-emergence chain~\eqref{eq:co_emergence} reads as a causal
sequence: mass ``produces'' signature, signature ``provides'' time, time
``enables'' Hilbert space. This temporal reading must be resisted. The
block does not first acquire mass, then a signature, then time, then
quantum mechanics. It satisfies all constraints at all levels
simultaneously. The chain is our explanation of the constraint structure,
expressed in the only order available to linear text, not a description
of how the block came to be.
\end{remark}

%======================================================================
\section{The self-consistency hierarchy}
\label{sec:hierarchy}
%======================================================================

We propose a four-level hierarchy of self-consistency constraints. The
levels are organized by \emph{coarse-graining}: Level~0 is the coarsest
description, Level~3 the finest accessible to a subsystem. The levels
are not temporal stages. They are simultaneous aspects of the same
four-dimensional block, described at different resolutions. A valid block
must satisfy the constraints at every level, and these constraints
interlock.

\begin{definition}[The four-level hierarchy]
\label{def:hierarchy}
\quad
\begin{itemize}
\item[\emph{Level 0.}] \emph{Topological.} The fundamental object is a
piecewise-linear (PL) 4-manifold $\mathcal{M}$, equipped with no
additional structure.

\item[\emph{Level 1.}] \emph{Smooth.} Given the topological type of
$\mathcal{M}$, the space of compatible smooth structures
$\mathrm{Sm}(\mathcal{M})$ is the configuration space of the theory. A
self-consistency measure $\mu^*$ on $\mathrm{Sm}(\mathcal{M})$ replaces
both the wavefunction and the path integral measure of standard
formulations.

\item[\emph{Level 2.}] \emph{Metric.} Each smooth structure constrains
the space of compatible metrics. The semiclassical Einstein
equation~\eqref{eq:sce} is the fixed-point condition that selects the
metric at this level; its existence is established in the companion
paper~\cite{fixed_point}.

\item[\emph{Level 3.}] \emph{Effective quantum mechanics.} Marginalizing
the smooth-structure measure $\mu^*$ over the complement of a subregion
$\mathcal{R} \subset \mathcal{M}$ produces an effective density matrix
$\rho_\mathcal{R}$ on a local Hilbert space
$\mathcal{H}_\mathcal{R}$. This level functions only for Lorentzian
signature with massive field content (Section~\ref{sec:coemergence}).
\end{itemize}
\end{definition}

The remainder of this section develops each level and specifies the
constraints that link them.

%----------------------------------------------------------------------
\subsection{Level 0: The topological constraint}
\label{sec:level0}
%----------------------------------------------------------------------

The manifold $\mathcal{M}$ must satisfy two requirements simultaneously:
it must admit a Lorentzian metric, and it must support nontrivial smooth
structures. These requirements are in tension.

A closed manifold admits a Lorentzian metric if and only if its Euler
characteristic vanishes: $\chi(\mathcal{M}) = 0$~\cite{geroch}. For
closed simply-connected 4-manifolds---the class where Freedman's
classification theorem~\cite{freedman} is strongest---the Euler
characteristic is $\chi = 2 + b_2$, where $b_2 = \mathrm{rank}(Q_\mathcal{M})$
is the second Betti number. Whenever the intersection form $Q_\mathcal{M}$ is
nontrivial ($b_2 \geq 1$), we have $\chi \geq 3$. The manifold class
where the Freedman--Donaldson machinery provides the richest structure
\emph{cannot carry a Lorentzian metric}.

This obstruction forces the Level~0 manifold away from the closed
simply-connected class. Three alternatives admit both $\chi = 0$ and
nontrivial smooth structures:

\begin{enumerate}
\item \textbf{Non-compact manifolds (exotic $\mathbb{R}^4$).} The
standard $\mathbb{R}^4$ admits uncountably many exotic smooth
structures~\cite{donaldson,taubes}. It has $\chi = 0$ and admits
Lorentzian metrics. However, $H_2(\mathbb{R}^4;\mathbb{Z}) = 0$, so
the intersection form is trivial. The rich topological structure is
carried by the smooth structures themselves, not by the homology.

\item \textbf{Manifolds with boundary (cobordisms).} Compact manifolds
with boundary can have $\chi = 0$ and nontrivial smooth structures. The
boundary data may carry topological information that compensates for the
loss of the closed-manifold intersection form.

\item \textbf{Non-simply-connected manifolds.} Manifolds with nontrivial
fundamental group (e.g., $T^4$) can have $\chi = 0$ and admit Lorentzian
metrics. The theory of exotic structures on non-simply-connected
4-manifolds is less developed than for simply-connected ones.
\end{enumerate}

Which class of 4-manifolds serves as the correct Level~0 input is an
open problem (Section~\ref{sec:open}). The co-emergence thesis does not
depend on the resolution: it requires only that $\mathcal{M}$ admits a
Lorentzian metric and supports enough smooth structures for Level~1 to
be nontrivial.

%----------------------------------------------------------------------
\subsection{Level 1: The smooth-structure measure}
\label{sec:level1}
%----------------------------------------------------------------------

The most remarkable fact in four-dimensional topology is the following:

\begin{theorem}[Donaldson, 1983; Taubes, 1987~\cite{donaldson,taubes}]
\emph{(Rigorous.)}
\label{thm:exotic}
The smooth manifold $\mathbb{R}^4$ admits uncountably many mutually
non-diffeomorphic smooth structures. In every other dimension $n \neq 4$,
$\mathbb{R}^n$ admits a unique smooth structure up to diffeomorphism.
\end{theorem}

The space $\mathrm{Sm}(\mathcal{M})$ of smooth structures compatible
with the PL type of a four-dimensional $\mathcal{M}$ is generically an
uncountable set. In any other dimension, it is a finite set (often a
single point).

\begin{definition}[The smooth-structure measure]
\emph{(Conjecture.)}
\label{def:mu}
The self-consistency measure $\mu^*$ on $\mathrm{Sm}(\mathcal{M})$ is
the unique measure satisfying the self-consistency condition
$\mathcal{F}(\mu^*) = \mu^*$, where $\mathcal{F}$ is the map defined
by: given a trial measure $\mu$ on $\mathrm{Sm}(\mathcal{M})$, compute
the induced metric and matter content at each smooth structure (via
Level~2), compute the resulting stress-energy, solve the Einstein
constraint for a new geometry, and update the weights on
$\mathrm{Sm}(\mathcal{M})$ accordingly.
\end{definition}

\begin{remark}[Relation to the path integral]
The smooth-structure measure $\mu^*$ is an analogue of the gravitational
path integral, but with a structural difference. The standard
gravitational path integral integrates over metrics---a continuous
infinite-dimensional space with an undefined measure. The measure $\mu^*$
integrates over smooth structures on a fixed topological type. While
$\mathrm{Sm}(\mathcal{M})$ may be uncountable, its structure as a moduli
space of gauge-equivalence classes is better controlled than the space of
all metrics. Donaldson and Seiberg--Witten theory provide invariants of
smooth structures that can serve as weights in $\mu^*$, even if the full
measure is not yet defined.
\end{remark}

%----------------------------------------------------------------------
\subsection{Level 2: The metric fixed point}
\label{sec:level2}
%----------------------------------------------------------------------

Given a smooth structure $\sigma \in \mathrm{Sm}(\mathcal{M})$, the
self-consistency constraint at the metric level is
eq.~\eqref{eq:sce}: the geometry that hosts the quantum state must be
sourced by that state. The companion paper~\cite{fixed_point} establishes
existence of solutions at three levels of rigor:

\begin{enumerate}

\item \emph{Exactly}, in the cosmological case via the Starobinsky
(1980) trace anomaly solution~\cite{starobinsky}, with
$H_0^2 = 180\pi/G|a_2|$ (Rigorous).

\item \emph{Perturbatively}, near any classical solution, via a Banach
contraction with constant $\kappa \sim (m/M_P)^2 \ll 1$ for massive
fields (Rigorous). The extension to massless fields remains open: the
spectral gap that controls the infrared behavior of the response kernel
vanishes at $m = 0$.

\item \emph{Conditionally non-perturbatively}, via the Schauder
fixed-point theorem on globally hyperbolic spacetimes with compact Cauchy
surfaces, under six stated assumptions of which four are established and
two remain open (Sketch).

\end{enumerate}

The Planck scale emerges as the natural validity boundary: when
$R \sim \ell_P^{-2}$, the contraction constant $\kappa \to 1$, the
perturbative expansion breaks down, and the full smooth-structure measure
$\mu^*$ becomes relevant.

\begin{remark}[Level~2 is signature-blind]
\label{rem:signature_blind}
The contraction constant $\kappa \sim (m/M_P)^2$ is identical for
Lorentzian and Riemannian metrics. The Banach contraction establishes
existence of self-consistent solutions regardless of signature. For
massless conformal matter, the Starobinsky de Sitter solution
(Lorentzian) and the round $S^4$ (Riemannian) are both exact
self-consistent fixed points~\cite{starobinsky,bunch_davies}. Level~2
does not distinguish signatures for any field content. This is not a
deficiency---it is informative: it demonstrates that signature selection
must operate at the cross-level constraint
(Conjecture~\ref{conj:coemergence}), not at any single level.
\end{remark}

%----------------------------------------------------------------------
\subsection{Level 3: Effective quantum mechanics}
\label{sec:level3}
%----------------------------------------------------------------------

\begin{conjecture}[Emergence of effective quantum mechanics]
\emph{(Conjecture.)}
\label{conj:qm}
Let $\mathcal{R} \subset \mathcal{M}$ be a subregion of the block. Let
$\mu^*$ be the self-consistency measure on $\mathrm{Sm}(\mathcal{M})$.
The marginal of $\mu^*$ over the smooth structures of the complement
$\mathcal{M} \setminus \mathcal{R}$, restricted to $\mathcal{R}$, defines
an effective measure $\mu^*_\mathcal{R}$ on the local configuration space
of $\mathcal{R}$.

When $\mathcal{R}$ is much larger than the Planck scale but much smaller
than the cosmological scale, and gravity within $\mathcal{R}$ is weak,
$\mu^*_\mathcal{R}$ is approximated by a density matrix
$\rho_\mathcal{R}$ on a Hilbert space $\mathcal{H}_\mathcal{R} =
L^2(\mathrm{Config}(\mathcal{R}), \mu^*_\mathcal{R})$.
\end{conjecture}

The Hilbert space $\mathcal{H}_\mathcal{R}$ is derived, not
postulated: it is the $L^2$ space of the marginal measure.
It is subsystem-dependent: different choices of $\mathcal{R}$ give
different Hilbert spaces. There is no single Hilbert space of the
universe. It is approximate: the density matrix approximation is valid
only in the sub-Planckian, weak-gravity regime.

This level functions only for Lorentzian signature with massive field
content. On a Riemannian manifold, the wave operator $\Delta + m^2$ is
elliptic: there are no propagating modes, no mass shell, no particle
interpretation, and no timelike geodesics to provide the local proper
time required for the Page--Wootters mechanism~\cite{page_wootters}. The
parameter $m^2$ is present in the field equation but is not
\emph{mass}: it generates no dispersion, no propagation, no local
clocks. For massless conformal matter on a Lorentzian background, proper
time exists along timelike geodesics, but without massive clock
subsystems the Page--Wootters conditioning has no temporal reference.
Level~3 collapses.

\begin{remark}[Why marginalization produces quantum statistics]
The argument for Conjecture~\ref{conj:qm} parallels the derivation of the
canonical ensemble from microcanonical statistics: integrating out the
environment imposes linearity on the subsystem. In the present context,
the consistency constraints on the full block, restricted to
$\mathcal{R}$, become approximately linear in the degrees of freedom of
$\mathcal{R}$ when the complement is integrated over, because the
dominant contribution from the complement is the mean-field background.
The analogy to statistical mechanics is structural---in both cases,
integrating out a large environment produces an approximately linear
effective theory---but the mechanism differs. Whether the marginalization
of $\mu^*$ produces the same linearization is the content of
Conjecture~\ref{conj:qm}, not a consequence of the analogy.
\end{remark}

%----------------------------------------------------------------------
\subsection{The interlocking constraint}
\label{sec:interlock}
%----------------------------------------------------------------------

The four levels are not independent. The constraint is that the
self-consistency chain must close:
\begin{equation}
\underbrace{\mathcal{M}}_{\text{Level 0}}
\xrightarrow{\text{smooth}}
\underbrace{\mathrm{Sm}(\mathcal{M})}_{\text{Level 1}}
\xrightarrow{\mu^* = \mathcal{F}(\mu^*)}
\underbrace{g^*}_{\text{Level 2}}
\xrightarrow{\text{Einstein}}
\underbrace{\langle\hat{T}_{\mu\nu}\rangle}_{\text{Level 2}}
\xrightarrow{\text{marginalize}}
\underbrace{\rho_\mathcal{R}}_{\text{Level 3}}
\xrightarrow{\text{consistency}}
\mathcal{M}
\label{eq:chain}
\end{equation}

The arrows represent logical constraints, not temporal sequences. A
block satisfying the hierarchy satisfies all constraints simultaneously.
The co-emergence thesis (Conjecture~\ref{conj:coemergence}) is the claim
that the joint solution space of this chain is restricted to Lorentzian
signature with massive field content: this is the only combination for
which all four levels are simultaneously nontrivial.

\begin{remark}[The chain is an epistemic decomposition]
The levels in~\eqref{eq:chain} are introduced for human
comprehensibility. They correspond to the precision at which we can
currently do mathematics. The block does not ``proceed'' through these
levels. It satisfies the constraint. The iteration suggested by
Axiom~\ref{ax:timeless} is our procedure for computing the fixed point,
not a description of how the universe came to be.
\end{remark}

%======================================================================
\section{Why four dimensions}
\label{sec:dimensions}
%======================================================================

The hierarchy of Section~\ref{sec:hierarchy} operates on a 4-manifold.
Here we argue that the four-dimensionality is not an input but a
consequence: four is the unique dimension where the constraint structure
of Definition~\ref{def:hierarchy} is nontrivial.

\begin{table}[h]
\centering
\begin{tabular}{lccc}
\toprule
Dimension & Smooth structures on $\mathbb{R}^n$ & $h$-cobordism theorem & Level~1 richness \\
\midrule
$n < 4$ & Unique & Holds (Smale, $n \geq 6$; special cases $n = 5$) & Insufficient \\
$n = 4$ & Uncountably many & \textbf{Fails} (Donaldson 1987) & Rich \\
$n > 4$ & Unique ($n \geq 5$) & Holds & Insufficient \\
\bottomrule
\end{tabular}
\caption{Topology by dimension. The failure of the $h$-cobordism theorem in
dimension four is the root of the exotic structure richness.}
\label{tab:dimensions}
\end{table}

\begin{proposition}[Four-dimensional uniqueness]
\emph{(Sketch.)}
\label{prop:4d}
The self-consistency hierarchy of Definition~\ref{def:hierarchy} is trivial
in every dimension other than four:
\begin{enumerate}
\item For $n \neq 4$: $\mathrm{Sm}(\mathcal{M})$ is a finite set
(generically a single point for $\mathbb{R}^n$). The Level~1 measure
$\mu^*$ is a finite sum, not an integral over a continuous moduli space.
The effective quantum mechanics of Conjecture~\ref{conj:qm} collapses to
classical statistics.

\item For $n = 4$: $\mathrm{Sm}(\mathcal{M})$ is uncountably infinite.
The Level~1 measure is a genuine integral over a continuous moduli space.
\end{enumerate}
\end{proposition}

\noindent\emph{Sketch.} Point~(1) uses Theorem~\ref{thm:exotic}: smooth
structures on $\mathbb{R}^n$ are unique for $n \neq 4$. The same
conclusion holds for compact manifolds in dimensions $n \geq 5$ (Smale's
$h$-cobordism theorem; see~\cite{smale}), and for $n \leq 3$ by
classical surface and 3-manifold theory. For $n = 4$,
Theorem~\ref{thm:exotic} gives the uncountable smooth structures.
The identification of the uncountable smooth structure moduli space as
the source of quantum statistics (Conjecture~\ref{conj:qm}) is a
conjecture, not a theorem. \hfill$\square$

\begin{remark}
Proposition~\ref{prop:4d} does not \emph{prove} that the universe is
four-dimensional. It establishes that the framework proposed here is
nontrivial only in dimension four. If the framework is correct,
four-dimensionality follows. Whether the framework is correct is an open
question.
\end{remark}

%======================================================================
\section{Connection to existing formulations}
\label{sec:connections}
%======================================================================

\subsection{The Wheeler--DeWitt equation as a Level-2 projection}

The Wheeler--DeWitt equation $\hat{H}\Psi[\gamma_{ij}] =
0$~\cite{dewitt} is the canonical quantization of the metric-level
constraint. In the hierarchy of Section~\ref{sec:hierarchy}, it is an
approximation to the Level~2 fixed-point condition, obtained by:

\begin{enumerate}
\item Concentrating the smooth-structure measure $\mu^*$ on a single
smooth structure (ignoring the Level~1 sum over exotic structures).
\item Performing a $3+1$ decomposition of the metric (introducing a
preferred foliation).
\item Replacing the Level~2 classical fixed-point condition with its
quantum analogue via Dirac quantization.
\end{enumerate}

Each step introduces additional structure not present in the hierarchy.
Step~(2) is the primary source of Axiom~\ref{ax:timeless}
violation~\cite{kuchar}. Steps~(1) and~(3) are simplifications that are
valid in the sub-Planckian, single-smooth-structure regime.

The Wheeler--DeWitt equation is thus the projection of the full hierarchy
onto a single smooth structure and a single spatial foliation. It
contains the correct physics at leading order in $\kappa \sim (m/M_P)^2$,
but not the global, foliation-independent constraint structure of the
full hierarchy.

\subsection{The semiclassical Einstein equation as Level-2}

The companion paper~\cite{fixed_point} studies eq.~\eqref{eq:sce} as a
fixed-point condition in its own right. Within the present hierarchy,
this is the Level~2 constraint evaluated on a single smooth structure.
The Banach contraction result ($\kappa \ll 1$ for sub-Planckian
curvatures) is the perturbative stability of the fixed point at this
level. The Schauder existence result (conditional on two open conjectures)
is non-perturbative existence.

The hierarchy extends this program: Level~2 existence (established with
caveats in the companion paper) is a necessary condition for the full
hierarchy to have a fixed point, but not sufficient. Level~1 requires
additionally that the measure over smooth structures has a fixed point.
Level~0 requires that the manifold type is self-consistent with the
matter content.

\subsection{The Page--Wootters mechanism at Level~3}

The Page--Wootters mechanism~\cite{page_wootters}---apparent time
evolution of a subsystem recovered from a globally static state via
conditioning on a ``clock'' subsystem---is a Level~3 phenomenon in the
present hierarchy. The globally static state is the smooth-structure
measure $\mu^*$; the conditioning on a subsystem is the marginalization
of Conjecture~\ref{conj:qm}. The apparent time evolution is a pattern
in the correlations between $\mathcal{R}$ and its complement encoded in
$\mu^*$.

The clock subsystem must be massive: it must experience proper time
along a timelike worldline (Section~\ref{sec:time_hilbert}). This
requirement connects Level~3 directly to the co-emergence thesis:
without mass, no clock; without a clock, no conditioning; without
conditioning, no effective time evolution and no local Hilbert space.

The observer-dependence identified by Kucha\v{r}~\cite{kuchar} and
confirmed by Marletto and Vedral~\cite{marletto_vedral}---different
choices of ``clock'' subsystem give different dynamics---is, in the
present hierarchy, a manifestation of the fact that different choices
of $\mathcal{R}$ give different marginal measures $\mu^*_\mathcal{R}$
and hence different effective Hilbert spaces
$\mathcal{H}_\mathcal{R}$. The observer-dependence is a feature:
it reflects the emergent, subsystem-relative nature of Hilbert space.

\subsection{Osterwalder--Schrader reconstruction as flat-space evidence}

The Osterwalder--Schrader (OS) reconstruction
theorem~\cite{osterwalder_schrader} provides independent evidence for the
mass-signature link: if a Euclidean quantum field theory has a mass gap
and satisfies reflection positivity, the theory reconstructs uniquely to
a Lorentzian QFT. The chain is: mass gap $\to$ reflection positivity
$\to$ Lorentzian theory. This is a rigorous result on flat space.

However, the OS theorem cannot serve as the \emph{mechanism} for
signature selection in the hierarchy, for two reasons. First, the OS
axioms require a flat Euclidean background with $E(4)$ invariance and a
global reflection hyperplane---fixed background structure that violates
Axiom~\ref{ax:nobackground}. Second, the reflection hyperplane defines a
preferred direction that becomes time after Wick rotation, which violates
Axiom~\ref{ax:timeless}.

The OS theorem is evidence, not mechanism: it demonstrates that the
connection between mass and Lorentzian signature is a rigorous
mathematical phenomenon, not verbal hand-waving. In the hierarchy, it
applies as a consistency check in the flat-space (weak-gravity) limit
of Level~2. The effective QFT obtained by marginalizing $\mu^*$ should
satisfy reflection positivity---a nontrivial constraint on the
self-consistency measure.

%======================================================================
\section{Open problems}
\label{sec:open}
%======================================================================

We list four priority open problems, in order of importance.

\begin{enumerate}

\item \textbf{The mass gap origin.}
The co-emergence thesis is conditional: \emph{if} mass, \emph{then}
signature, time, and Hilbert space co-emerge. Where does mass come from?
The hierarchy's proven results do not depend on answering this question---they
hold for any mass input satisfying the perturbative bound $m \ll M_P$.
But the framework aspires to derive mass from topology
(Axiom~\ref{ax:nobackground}).

Mass enters the hierarchy at Level~2 as a parameter of the matter field
content. Cross-level consistency requires at least some massive fields
(Conjecture~\ref{conj:coemergence}): without mass, there are no timelike
geodesics, no local proper time, no local Hilbert space, and therefore
no Level~3 structure. Whether the mass spectrum can be derived from
Levels~0--1 is the hierarchy's deepest open problem.

\item \textbf{Level~3 formalization.}
Make ``local Hilbert space requires local time requires mass''
mathematically precise. The Page--Wootters mechanism provides a partial
model, but as standardly formulated it presupposes a total Hilbert
space (Remark~\ref{rem:pw_gap}). The needed result: reformulate
Page--Wootters using the marginal of $\mu^*$ rather than a fundamental
Hilbert space, and show that the conditioning on a massive clock
subsystem requires the metric to be Lorentzian. This would promote the
co-emergence thesis from Conjecture to Sketch.

\item \textbf{The Level~0 manifold class.}
Which 4-manifolds admit both Lorentzian metrics ($\chi = 0$) and
nontrivial smooth structures? The Euler characteristic obstruction
(Section~\ref{sec:level0}) rules out closed simply-connected manifolds
with nontrivial intersection forms. The candidates---exotic
$\mathbb{R}^4$, manifolds with boundary, non-simply-connected
manifolds---each have different mathematical properties. Identifying the
correct class is prerequisite to making Level~0 precise.

\item \textbf{A natural measure on smooth structures.}
Define a measure $\mu$ on $\mathrm{Sm}(\mathcal{M})$ for a 4-manifold
$\mathcal{M}$ that is:
\begin{enumerate}
\item Defined without reference to a metric (to avoid the circularity of
the standard gravitational measure problem).
\item Compatible with the self-consistency condition
$\mathcal{F}(\mu) = \mu$.
\item Invariant under the orientation-preserving diffeomorphism group.
\end{enumerate}
Donaldson--Witten theory and Seiberg--Witten invariants provide natural
weights on smooth structures; assembling these into a well-defined
measure on the full moduli space $\mathrm{Sm}(\mathcal{M})$ is the open
problem.

\end{enumerate}

%======================================================================
\section*{Acknowledgments}
%======================================================================

% [placeholder]

\begin{thebibliography}{99}

\bibitem{fixed_point}
W.~Regelmann and Claude (Anthropic),
``Fixed-point existence of self-consistent semiclassical gravity,''
companion paper (2026).

\bibitem{dewitt}
B.~S.~DeWitt,
``Quantum theory of gravity. I. The canonical theory,''
\textit{Phys.\ Rev.} \textbf{160}, 1113 (1967).

\bibitem{wheeler}
J.~A.~Wheeler,
``Superspace and the nature of quantum geometrodynamics,''
in \textit{Battelle Rencontres}, eds.\ C.~M.~DeWitt and J.~A.~Wheeler
(Benjamin, New York, 1968), pp.~242--307.

\bibitem{kuchar}
K.~V.~Kucha\v{r},
``Time and interpretations of quantum gravity,''
in \textit{Proc.\ 4th Canadian Conference on General Relativity and
Relativistic Astrophysics}, eds.\ G.~Kunstatter, D.~Vincent, and
J.~Williams (World Scientific, 1992), pp.~211--314.

\bibitem{page_wootters}
D.~N.~Page and W.~K.~Wootters,
``Evolution without evolution: Dynamics described by stationary observables,''
\textit{Phys.\ Rev.\ D} \textbf{27}, 2885 (1983).

\bibitem{wigner}
E.~P.~Wigner,
``On unitary representations of the inhomogeneous Lorentz group,''
\textit{Ann.\ Math.} \textbf{40}, 149--204 (1939).

\bibitem{geroch}
R.~Geroch,
``Spinor structure of space-times in general relativity.\ I,''
\textit{J.\ Math.\ Phys.} \textbf{9}, 1739--1744 (1968).

\bibitem{starobinsky}
A.~A.~Starobinsky,
``A new type of isotropic cosmological models without singularity,''
\textit{Phys.\ Lett.\ B} \textbf{91}, 99--102 (1980).

\bibitem{bunch_davies}
T.~S.~Bunch and P.~C.~W.~Davies,
``Quantum field theory in de Sitter space: Renormalization by point-splitting,''
\textit{Proc.\ R.\ Soc.\ Lond.\ A} \textbf{360}, 117--134 (1978).

\bibitem{freedman}
M.~H.~Freedman,
``The topology of four-dimensional manifolds,''
\textit{J.\ Diff.\ Geom.} \textbf{17}, 357--453 (1982).

\bibitem{donaldson}
S.~K.~Donaldson,
``An application of gauge theory to four-dimensional topology,''
\textit{J.\ Diff.\ Geom.} \textbf{18}, 279--315 (1983).

\bibitem{smale}
S.~Smale,
``On the structure of manifolds,''
\textit{Amer.\ J.\ Math.} \textbf{84}, 387--399 (1962).

\bibitem{taubes}
C.~H.~Taubes,
``Gauge theory on asymptotically periodic $4$-manifolds,''
\textit{J.\ Diff.\ Geom.} \textbf{25}, 363--430 (1987).

\bibitem{hawking_ellis}
S.~W.~Hawking and G.~F.~R.~Ellis,
\textit{The Large Scale Structure of Space-Time}
(Cambridge University Press, 1973).

\bibitem{halliwell}
J.~J.~Halliwell,
``Derivation of the Wheeler--DeWitt equation from a path integral for
minisuperspace models,''
\textit{Phys.\ Rev.\ D} \textbf{38}, 2468 (1988).

\bibitem{banks}
T.~Banks,
``TCP, quantum gravity, the cosmological constant and all that\ldots,''
\textit{Nucl.\ Phys.\ B} \textbf{249}, 332 (1985).

\bibitem{marletto_vedral}
C.~Marletto and V.~Vedral,
``Evolution without evolution and without ambiguities,''
\textit{Phys.\ Rev.\ D} \textbf{95}, 043510 (2017).

\bibitem{osterwalder_schrader}
K.~Osterwalder and R.~Schrader,
``Axioms for Euclidean Green's functions,''
\textit{Commun.\ Math.\ Phys.} \textbf{31}, 83--112 (1973);
``Axioms for Euclidean Green's functions.\ II,''
\textit{Commun.\ Math.\ Phys.} \textbf{42}, 281--305 (1975).

\end{thebibliography}

\end{document}
